---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data

Setting "echo=TRUE" for all code chunks
```{r Setting_Global_Options,echo=TRUE}
library(knitr)
opts_chunk$set(echo=TRUE)
```

### Data Loading
```{r Data_Loading}
setwd("C:/Users/user/Desktop/Data Analytics Stuff/Coursera/Reproducible research")
activity<-read.csv("activity.csv",header=TRUE)
```

### Data Processing
Removing NAs
```{r Data_Processing}
act<-!(is.na(activity))
activity0<-activity[act,c(1:3)]
str(activity0)
head(activity0)
```

## Mean of Total Number of steps taken per day

### Total Number of steps taken per day
```{r Total_no_of_Steps_per_day}
library(dplyr)
activity1<-group_by(activity0,date) %>% summarize(total_steps=sum(steps,na.rm = TRUE))
head(activity1)
```

### Histogram of the total number of steps taken each day
```{r Histogram}
hist(activity1$total_steps,breaks=20,xlab = "Total Number of Steps taken per day",main="Histogram")
```

### Mean and Median of total number of steps taken per day
```{r Mean_and_Median}
mean(activity1$total_steps)
median(activity1$total_steps)
```

## Average daily activity pattern

### 1.Time Series Plot(With original data)
```{r plot}
activity2<-group_by(activity0,interval) %>% summarize(avg_no_of_steps=mean(steps))
plot(activity2$interval,activity2$avg_no_of_steps,xlab="Interval",ylab="Avg Number of steps",type="l")
```

### 2.Interval which contains on average Maximum Number of steps across all the days.
```{r maximum_point_in_graph}
plot(activity2$interval,activity2$avg_no_of_steps,xlab="Interval",ylab="Avg Number of steps",type="l")
abline(v=activity2$interval[which.max(activity2$avg_no_of_steps)])
```

## Imputing Missing Values

### 1.Total Number of Missing Values in the Dataset
```{r Total_NAs}
sum(is.na(activity))
```

### 2.The strategy for filling in all of the missing values in the dataset

Mean for 5-minute interval is used to fill in the missing values in the dataset

```{r Imputing_NAs}
head(activity2)
```

### 3.The New imputed Dataset
```{r Imputing_Missing_Values}
activity3<-activity
for(i in 1:nrow(activity3)){
  if(is.na(activity3$steps[i])){
    activity3$steps[i]<-activity2$avg_no_of_steps[which(activity2$interval==activity3$interval[i])]
  }
}
head(activity3)
```


### 4.Histogram(Imputed Dataset) of the total number of steps taken each day
```{r Histogram_of_total_number_of_steps_taken_each_day}
library(dplyr)
activity4<-group_by(activity3,date) %>% summarize(total_steps=sum(steps))
head(activity4)
hist(activity4$total_steps,breaks=20,xlab = "Total Number of Steps taken per day",main="Histogram of the Imputed Data")
```

### Mean and Median(Imputed Dataset) of total number of steps taken per day
```{r Mean_and_Median(Imputed_Dataset)}
mean(activity4$total_steps)
median(activity4$total_steps)
```

### Comparison between Dataset with missing Values and Dataset with Imputed Data.
```{r Comparison_between_means}
barplot(c(mean(activity1$total_steps),mean(activity4$total_steps)),names.arg=c("Mean of total steps\n(With NAs)","Mean of total steps\n(Imputed data)"),main="Comparison between means of two Datasets",ylim=c(10400,11000))
```

```{r Comparison_between_medians}
barplot(c(median(activity1$total_steps),mean(activity4$total_steps)),main="Comparison between medians of two Datasets",ylim=c(10600,10900),names.arg=c("Median of total steps\n(With NAs)","Median of total steps\n(Imputed data)"))
```


## To check Differences in activity patterns between weekdays and weekends.

### 1.New factor variable in the dataset with two levels - "weekday" and "weekend"
```{r dayType}
library(chron)
for(i in 1:nrow(activity3)){
  if(is.weekend(activity3$date[i])){
    activity3$daytype[i]="Weekend"
  }
  else{
    activity3$daytype[i]="Weekday"
  }
}
activity3$daytype<-as.factor(activity3$daytype)
str(activity3)
```

### 2.Time Series Plot(on the basis of Day Type)
```{r time_series_plot(on_the_basis_of_Day_Type)}
activity5<-group_by(activity3,interval,daytype) %>% summarize(avg_no_of_steps=mean(steps))
library(ggplot2)
g<-ggplot(activity5,aes(interval,avg_no_of_steps))
g+geom_line()+facet_grid(daytype~.)+xlab("Interval")+ylab("Number of steps")
```
